---
title: "Births in Kansas (1991)"
author: "Sodahm Robin Yoo"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: html_document
---

<!--
	This code chunk suppresses all code chunks from appearing in the rendered
  document
-->

```{r, suppress-code}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```

<!--
	The code chunk below reads in the data from my local drive.
-->

```{r, load-data}

load("birth.Rdata")

```

<!--
  The code chunk below loads in the package s, which you'll have to install if you haven't already
-->

```{r, required-pckgs}

library(tidyverse)

```

<!--
	The code chunk below is a function to create a unique ID for each record.
-->

```{r, unique-ID}

## Function to create unique IDs
create_unique_ids <- function(n, seed_no = 1, char_len = 5) {

  set.seed(seed_no)
  pool <- c(letters, LETTERS, 0:9)

  res <- character(n)  # pre-allocating vector is much faster than growing it
  for (i in seq(n)) {
    this_res <- paste0(sample(pool, char_len, replace = TRUE), collapse = "")
    while (this_res %in% res) {
      # if there was a duplicate, redo
      this_res <- paste0(sample(pool, char_len, replace = TRUE), collapse = "")
    }
    res[i] <- this_res
  }
  res
}

```

<!--
  The code chunk below does some cleaning for my birth records datasets
-->

```{r, data-cleaning}

id <- create_unique_ids(nrow(birth))
birth$id <- id

vars <- c("id", "dmage", "mrace3", "dmeduc", "dmar", "nprevis", "birmon", "biryr", "weekday", "dgestat",
"csex", "dbirwt", "dplural", "diabetes", "chyper", "preterm", "tobacco", "alcohol", "congan")

## dataset to be cleaned for plausible values
birth <- birth[vars]

birth <- mutate(birth,
  ID = id,
  AGE_M = ifelse(dmage >= 10 & dmage <= 55, dmage, NA),
  RACE_M = case_when(mrace3 == "1" ~ "1", ## white
  mrace3 == "2" ~ "1", ## black
  mrace3 == "3" ~ "3"), ## other
  MARITAL = case_when(dmar == "1" ~ "1", ## married
  dmar == "2" ~ "2"), ## unmarried
  PRENATAL = ifelse(nprevis == 99, NA, nprevis),
  BIRTHMONTH = as.character(birmon),
  BIRTHYEAR = as.character(biryr),
  BIRTHDAY = as.character(weekday),
  GEST_AGEDELIVER = ifelse(dgestat >= 20 & dgestat <= 40, dgestat, NA),
  SEX = case_when(csex == "1" ~ "1", ## male
  csex == "2" ~ "2"), ## female
  BIRTHWEIGHT = ifelse(dbirwt >= 250 & dbirwt <= 7500, dbirwt, NA),
  PLUR = ifelse(dplural > 0 & dplural < 6, dplural, NA),
  PREPREG_DIAB = as.character(diabetes),
  HTN = as.character(chyper),
  PREV_PRETERM = as.character(preterm),
  TOBACCO = as.character(tobacco),
  ALCOHOL = as.character(alcohol),
  CONANOM = as.character(congan),
  PTB = ifelse(GEST_AGEDELIVER <= 36, "1", "0"),
  LBW = ifelse(BIRTHWEIGHT < 2500, "1", "0")) %>%
  select(-vars)

```
## Executive summary

These birth records are for the state of Kansas in 1991. We are interested in
examining the associations between the covariates in the data set and adverse
birth outcomes, including low birthweight.

There are `r nrow(birth)` birth records in the data set, with `r ncol(birth)`
variables for each record.

Low birthweight is defined as a weight less than 2500 grams, regardless of gestational age.

Of `r nrow(birth)` births in Kansas in 1991, `r sum(birth$LBW == "1", na.rm = TRUE)` resulted in
low birthweights.

Overall, we found a `r ifelse(cor(birth$BIRTHWEIGHT, birth$AGE_M, use = "complete.obs") > 0, "positive", "negative")`
relationship between maternal age and birthweight.

## Analysis

We analyzed the relationship between maternal age (`AGE_M`) and birthweight (`BIRTHWEIGHT`).
A scatterplot is shown below.

```{r, make-scatter}

plot(birth$BIRTHWEIGHT, birth$AGE_M,
     xlab = "Birthweight (g)",
     ylab = "Maternal age (yrs)",
     pch = 19)

```

<!--
  Rscript -e "rmarkdown::render('info550-hw3.Rmd', 'ioslides_presentation', 'hw3.html')"
-->

<!--
  Participation grades:
  
  Nanxi: 1
  Limeng: 1
  Shauna: 1
-->
